---
title: "Pathing"
author: "Justina Zou"
date: "July 23, 2018"
output: html_document
---

This file differs from singleFileFromTo in that it's less interactive, but it is able to display more data in the form of tables. In addition, rather than just displaying location to location information, it also shows ap to ap information.

The goal is to find areas of underutilization by finding which paths are most common. Essentially, if there are many paths that start from location A and end in location B but don't go through the locations between A and B or, in the case that A and B are connected, if it goes around A and B instead of to A and B directly.

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
library(readr) # to read in csv files
library(dplyr) # to process/clean data
library(ggplot2) # to plot
library(sp) # for spatial polygons
library(rgdal)
library(lubridate)
library(raster)
library(deldir)
library(geosphere)
library(rgeos) 
library(data.table) # for creating column indicating same location
library(igraph)
coord <- read_csv("locationsToCoordinates.csv") # for location data
df0419 <- read_csv("mergedData0419.csv") # dataset we are using
```

## Modifying findIndex and howLong

This is done because I had trouble passing in parameters to the original functions and I figured out how to make them more efficient.

```{r}
# returns dataframe with columns
# macaddr, id, fromLoc, totalTime, startTime, and endTime
# where total time is how long the mac spent at the location
howLong <- function(macdf) {
  macdf <- macdf %>% # creating column with index that changes when location changes
    group_by(macaddr) %>% 
    mutate(id = rleid(fromLoc))
  macdf <- macdf %>% # summing up the amount of time spent at a location
    group_by(macaddr, id, fromLoc) %>% 
    summarise(totalTime = sum(timeDiff), 
              startTime = min(`_time`), 
              endTime = max(`_time`))
  return(macdf)
}

# Returns a list of two things
# 1. Original dataframe filtered down to relevant info
# 2. Truncated dataframe filtered to qualifying locations
findIndex <- function(macdf, fromLoc, toLoc, 
                      fromInte = 60 * 5, toInte = 60 * 5, 
                      betweenInte = 60 * 10, distInte = 2) {
  orig <- macdf # saving original dataframe
  macdf <- howLong(macdf)
  # filter dataframe based on parameters
  # e.g. whether the macaddr stayed in a location for the correct amount of time
  macdf <- macdf %>% mutate(doesQualify = (fromLoc == fromLoc & totalTime > fromInte) | 
                              (fromLoc == toLoc & totalTime > toInte) |
                              (!fromLoc %in% c(fromLoc, toLoc) & totalTime > betweenInte))
  macdf <- macdf[macdf$doesQualify, ]
  
  macdf <- macdf[!is.na(macdf$doesQualify), ] # getting rid of NA
  
  # finding relevant path from A to B
  macs <- unique(macdf$macaddr)
  ret <- NULL
  ret2 <- NULL
  sapply(macs, function(mac) {
    macLoc <- macdf %>% 
      filter(macaddr == mac)
    orig <- orig %>% 
      filter(macaddr == mac)
    
    if(!fromLoc %in% macLoc$fromLoc | !toLoc %in% macLoc$fromLoc) { # does not visit locations correctly
      return(NULL)
    }
    # finding when device visited locations
    indexFrom <- match(fromLoc, macLoc$fromLoc)
    indexTo <- match(toLoc, macLoc$fromLoc)
    
    while(!is.na(indexTo) & !is.na(indexFrom)) {
      if(indexTo > indexFrom) {
        # to minimize the distance between the two locations -- that is, for ex,
        # to filter out if a device starts at perkins, chills in their room
        # for two hours, goes on East, goes back to west, goes to WU.
        if(indexTo - indexFrom > distInte) {
          indexFrom <- match(fromLoc, macLoc[indexFrom+1:length(macLoc), ]$fromLoc) + indexFrom
          next
        }
        break
      }
      indexTo <- match(toLoc, macLoc[indexTo+1:length(macLoc), ]$fromLoc) + indexTo
    }
    # macaddr does not visit locations correctly
    if(is.na(indexTo) | is.na(indexFrom)) {
      return(NULL)
    }
    
    indexFrom2 <- match(macLoc$startTime[indexFrom], orig$`_time`)
    indexTo2 <- match(macLoc$startTime[indexTo], orig$`_time`)
    ret <<- rbind(ret, orig[indexFrom2:indexTo2, ])
    ret2 <<- rbind(ret2, macLoc[indexFrom:indexTo, ])
    return(NULL)
  })
  return(list(orig = ret, trun = ret2))
}
```

## Creating function findPaths

This function takes in the same parameters as findIndex, but instead it calls findIndex and returns three things.
1. The original return of findIndex - for debugging purposes.
2. The most popular complete paths.
3. The most popular, single location to location paths.
This is useful when one wants to visualize several different paths from different locations.
The dataframe, macdf, has 19 parameters.
prefix, ap, _time, asa_code, macaddr, slotnum, ssid, ipaddr, location.x, fromLoc, lat, long, campus, APname, APnum, org, toLoc, nextTime, timeDiff
fromLoc - location.y from mergedData
toLoc - next location visited by that macaddr
nextTime - the time of the next event registered by that macaddr
timeDiff - nextTime minus _time

```{r}
findPaths <- function(macdf, fromLoc, toLoc, 
                      fromInte = 60 * 5, toInte = 60 * 5, 
                      betweenInte = 60 * 10, distInte = 2) {
  findI <- findIndex(macdf, fromLoc, toLoc, 
                      fromInte, toInte, 
                      betweenInte, distInte)
  dfLocs <- findI$orig
  
  # Condensing for better pathing
  paths <- dfLocs %>% 
    group_by(macaddr) %>% 
    mutate(id = rleid(fromLoc))
  paths <- paths %>% # summing up the amount of time spent at a location
    group_by(macaddr, id, fromLoc) %>% 
    summarise(totalTime = sum(timeDiff)) 
  paths <- paths %>% 
    group_by(macaddr) %>% 
    mutate(toLoc = lead(fromLoc, order_by = macaddr))
  
  # Most common A -> B paths, that is, the full path rather than just the most popular destinations
  # E.g. shows A to C to D to B rather than
  # A to C, then C to D, then D to B, the latter of which is a "location to location" path
  macs <- unique(paths$macaddr)
  fullPaths <- list(path = NULL, count = NULL)
  # Count the frequencies of certain paths
  delete <- sapply(macs, function(mac) { # assigning to variable to suppress output
    
    macLoc <- paths %>%
      filter(macaddr == mac)
    path <- toString(paste0(macLoc$fromLoc))
    if(!path %in% fullPaths$path) { # path is not in paths list
      fullPaths$path <<- c(fullPaths$path, path) # add to list and initialize count
      fullPaths$count <<- c(fullPaths$count, 0)
    }
    index <- match(path, fullPaths$path)
    fullPaths$count[index] <<- fullPaths$count[index] + 1 # increment by one
    })
  
  fullPaths <- as.data.frame(fullPaths)
  fullPaths <- fullPaths %>%   
    arrange(desc(count))
  
  # Most common location to location paths
  paths <- paths %>% 
    group_by(fromLoc, toLoc) %>% 
    summarise(freq = n()) %>% 
    arrange(desc(freq))
  paths <- paths[!is.na(paths$toLoc), ]
  
  ret <- list(orig = findI, allPaths = fullPaths, subPaths = paths)
  
  return(ret)
}
```

## Creating the adjacency table

This is a table that shows which polygons are adjacent to each other. This is useful for comparing whether macs tend to go from adj polygon to adj polygon or they tend to skip neighboring polygons.

```{r adjtable, message = FALSE}
# calculating voronoi cells and converting to polygons to plot on map
z <- deldir(coord$long, coord$lat) # computes cells
# convert cell info to spatial data frame (polygons)
w <- tile.list(z)
polys <- vector(mode="list", length=length(w))
for (i in seq(along=polys)) {
  pcrds <- cbind(w[[i]]$x, w[[i]]$y)
  pcrds <- rbind(pcrds, pcrds[1,])
  polys[[i]] <- Polygons(list(Polygon(pcrds)), ID=as.character(i))
}
SP <- SpatialPolygons(polys)
for(x in 1:nrow(coord)){
  SP@polygons[[x]]@ID <- as.character(x)
}
SPDF <- SpatialPolygonsDataFrame(SP, data=data.frame(x=coord[,2], y=coord[,3]))
# tag polygons with location name
SPDF@data$ID = coord$location
a <- sapply(1:length(coord$location), function(x){
  SPDF@polygons[[x]]@ID <- coord$location[x]
  SPDF <<- SPDF
})
adjTable <- gTouches(SPDF, byid = TRUE)
```

## Initial filtering 

Filtering for phones because they're a better indicator of movement than laptops for the purposes of location to location pathing.
For general pathing, it may also be of interest to limit it to only phones, but because all the paths are wanted, and it's hard to tell which orgs are phones only, the general pathing is done using the full dataset.

```{r filter}
phones <- c("HTC Corporation", "zte corporation", 
            "Motorola Mobility LLC, a Lenovo Company",
            "LG Electronics Mobile Communications", 
            "TCT mobile ltd", "Sony Mobile Communications Inc")
dfphones <- df0419 %>% 
  filter(org %in% phones) %>% 
  arrange(`_time`, macaddr)
dfall <- df0419 %>% 
  arrange(`_time`, macaddr)

# Creating empty path matrix -- currently unused
pathsTable <- matrix(c(0), nrow = length(coord$lat), ncol = length(coord$long))
dimnames(pathsTable) <- list(c(coord$location), c(coord$location))

# Calculating next location and the time difference between
dfphones <- dfphones %>% # for location to location
  group_by(macaddr) %>% 
  mutate(toLoc = lead(location.y, order_by = macaddr),
         nextTime = lead(`_time`, order_by = macaddr),
         timeDiff = as.numeric(difftime(nextTime, `_time`, units = "secs"))) %>% 
  rename(fromLoc = location.y) 
dfallap <- dfall %>% # for most common paths between aps
  group_by(macaddr) %>% 
  mutate(toAP = lead(ap, order_by = macaddr),
         toLoc = lead(location.y, order_by = macaddr),
         nextTime = lead(`_time`, order_by = macaddr),
         timeDiff = as.numeric(difftime(nextTime, `_time`, units = "secs"))) %>% 
  rename(fromAP = ap,
         fromLoc = location.y) 
dfallloc <- dfall %>% # for most common paths between locations
  group_by(macaddr) %>% 
  mutate(toLoc = lead(location.y, order_by = macaddr),
         nextTime = lead(`_time`, order_by = macaddr),
         timeDiff = as.numeric(difftime(nextTime, `_time`, units = "secs"))) %>% 
  rename(fromLoc = location.y) 

# For general pathing purposes
dfGenLoc <- dfallloc %>% 
  dplyr::select(`_time`, nextTime, fromLoc, toLoc, macaddr, timeDiff)
dfGenLoc <- dfGenLoc[which(dfGenLoc$fromLoc != dfGenLoc$toLoc), ] # getting rid of A -> A
dfGenAP <- dfallap[which(dfallap$fromAP != dfallap$toAP), ]
```

## Finding paths

### Filtering for phones

This is finding paths by figuring out how a macaddr moves throughout the day.

```{r phones, message = FALSE, warning = FALSE, results = "asis"}
# For location to location pathing
# Parameters
fl <- c("Kilgo", "Keohane", "Perkins")
tl <- c("Perkins", "Perkins", "WestUnion")
fromInte <- c(60 * 5, 60 * 5, 60 * 5)
toInte <- c(60 * 5, 60 * 5, 60 * 5)
betweenInte <- c(60 * 5, 60 * 5, 60 * 5)
distInte <- c(1, 1, 1)

fullPaths <- vector("list", length(fl))
subPaths <- vector("list", length(fl))

for(i in 1:length(fl)) {
  # Filter for macs that have visited both locations to speed up runtimes
  bothMacs <- unique((dfphones %>%
                        filter(fromLoc %in% c(fl[i],tl[i])))$macaddr)
  dfphones <- dfphones %>% 
    filter(macaddr %in% bothMacs)
  
  # Finding viable paths
  temp <- findPaths(dfphones, fl[i], tl[i],
                    fromInte[i], toInte[i], betweenInte[i], distInte[i])
  paths <- temp$subPaths
  
  # Saving dataframes
  fullPaths[[i]] <- temp$allPaths
  subPaths[[i]] <- paths
  
  # Note that the paths may seem weird because of the particular orgs chosen to be represented.
  print(knitr::kable(temp$allPaths, format = "html"))
  
  print(knitr::kable(paths, format = "html"))
  
  # Viewing graph
  coordTrun <- coord %>% # getting rid of vertices that don't appear on graph
    filter(location %in% c(paths$fromLoc, paths$toLoc))
  coordMx <- as.matrix(coordTrun[, 3:2]) # placing vertices on geographic location
  g <- graph.data.frame(paths, directed = TRUE, vertices = coordTrun)
  E(g)$width <- sqrt(paths$freq) # wider == more frequent; sqrt size for easy viewings
  plot(g, 
       layout = coordMx, 
       edge.arrow.size = 0.15, 
       vertex.size = 1, 
       vertex.label.dist = 1, 
       vertex.color = "darkblue", 
       edge.color = "violet",
       vertex.label.cex = 0.75, 
       vertex.label.family = "Helvetica",
       vertex.label.font = 2, 
       main = paste("From", fl[i], "to", tl[i])) 
}
```

### Without filtering for phones

```{r nophones, message = FALSE, warning = FALSE, results = "asis"}
# For location to location pathing
# Parameters
fl <- c("Kilgo", "Keohane", "Perkins")
tl <- c("Perkins", "Perkins", "WestUnion")
fromInte <- c(60 * 5, 60 * 5, 60 * 5)
toInte <- c(60 * 5, 60 * 5, 60 * 5)
betweenInte <- c(60 * 5, 60 * 5, 60 * 5)
distInte <- c(1, 1, 1)

temp <- NULL

for(i in 1:length(fl)) {
  # Filter for macs that have visited both locations to speed up runtimes
  bothMacs <- unique((dfallloc %>%
                        filter(fromLoc %in% c(fl[i],tl[i])))$macaddr)
  bothMacs <- sample(bothMacs, min(500, length(bothMacs))) # sampling only a portion
  dfallFilt <- dfallloc %>% 
    filter(macaddr %in% bothMacs)
  
  # Finding viable paths
  temp <- findPaths(dfallFilt, fl[i], tl[i],
                    fromInte[i], toInte[i], betweenInte[i], distInte[i])
  paths <- temp$subPaths
  
  # Note that the paths may seem weird because of the particular orgs chosen to be represented.
  print(knitr::kable(temp$allPaths, format = "html"))
  
  print(knitr::kable(paths, format = "html"))
  
  # Viewing graph
  coordTrun <- coord %>% # getting rid of vertices that don't appear on graph
    filter(location %in% c(paths$fromLoc, paths$toLoc))
  coordMx <- as.matrix(coordTrun[, 3:2]) # placing vertices on geographic location
  g <- graph.data.frame(paths, directed = TRUE, vertices = coordTrun)
  E(g)$width <- sqrt(paths$freq) # wider == more frequent; sqrt size for easy viewings
  plot(g, 
       layout = coordMx, 
       edge.arrow.size = 0.15, 
       vertex.size = 1, 
       vertex.label.dist = 1, 
       vertex.color = "darkblue", 
       edge.color = "violet",
       vertex.label.cex = 0.75, 
       vertex.label.family = "Helvetica",
       vertex.label.font = 2, 
       main = paste("From", fl[i], "to", tl[i])) 
}
```

## Viewing the most popular paths in general

### By location

```{r genpaths}
dfGenTrun <- dfGenLoc %>% 
       group_by(fromLoc, toLoc) %>% 
       summarise(freq = n()) %>% 
       arrange(desc(freq))

# I have commented out this chunk because it can be useful to view all the possible connections, 
# but it may be of bigger interest to see if there's a lot more movement from Perk to Rub 
# but not vice-versa.
# # Merging A -> B and B -> A
# dfGenTrun <- graph.data.frame(dfGenTrun, directed = TRUE)
# dfGenTrun <- as.undirected(dfGenTrun, mode = "collapse", edge.attr.comb = "sum")
# dfGenTrun <- data.frame(get.edgelist(dfGenTrun),
#                   get.edge.attribute(dfGenTrun, "freq"))
# colnames(dfGenTrun) <- c("fromLoc", "toLoc", "freq")
dfGenTrun <- dfGenTrun %>% 
  arrange(desc(freq))

knitr::kable(dfGenTrun[1:20, ], format = "html")
```

### By AP

```{r}
dfGenAPTrun <- dfGenAP %>% 
  group_by(fromAP, toAP, fromLoc, toLoc) %>%
  summarise(freq = n()) %>% 
  arrange(desc(freq))

knitr::kable(dfGenAPTrun[1:20, ], format = "html")
```

## Observing paths that aren't adjacent

This could be an indication that more aps are needed, people move too fast to be caught by aps, the polygons drawn aren't representative, etc.

### By location

```{r notadjloc}
notAdjLoc <- NULL
delete <- NULL
delete <- sapply(1:length(dfGenTrun$fromLoc), function(i){
  from <- as.character(dfGenTrun$fromLoc[i])
  to <- as.character(dfGenTrun$toLoc[i])
  if(!adjTable[from, to]) { # if A and B aren't adjacent
    notAdjLoc <<- rbind(notAdjLoc, dfGenTrun[i, ]) # add to dataframe for further analysis
  }
})

knitr::kable(notAdjLoc[1:20, ], format = "html")
```

### By AP

```{r notadjap}
notAdjAP <- NULL
delete <- NULL
delete <- sapply(1:500, function(i){ # 1:length(dfGenAPTrun$fromLoc) ; using 100 to speed stuff up
  from <- as.character(dfGenAPTrun$fromLoc[i])
  to <- as.character(dfGenAPTrun$toLoc[i])
  if(!adjTable[from, to] & from != to) { # if A and B aren't adjacent
    notAdjAP <<- rbind(notAdjAP, dfGenAPTrun[i, ]) # add to dataframe for further analysis
  }
})

knitr::kable(notAdjAP[1:20, ], format = "html")
```

## Observing different location AP to AP jumps

By conducting this analysis we are hoping to discover if the data/calculations accurately reflect which buildings are physically next to each other as well as APs that could interfere with each others' signals. In addition, this is also a rudimentary way of finding out where the APs are, physically -- in this case, these APs are presumably near the exterior of buildings.

```{r}
locJumps <- dfGenAP[which(dfGenAP$fromLoc != dfGenAP$toLoc), ] # jumps where locations are different

locJumps <- locJumps %>% 
  group_by(fromAP, toAP, fromLoc, toLoc) %>%
  summarise(freq = n()) %>% 
  arrange(desc(freq))

knitr::kable(locJumps[1:20, ], format = "html")
```

## AP to AP pathing by unique occurences per mac

Only AP to AP pathing is further calculated by unique occurences per mac because one macaddr constantly switching between two APs can cause a unrepresentative number of events for those two APs while with location, it is less likely. 
Further observing locations by unique occurrences can possibly reveal discreptancies caused by problematic APs that are near the outside of the building.

```{r}
# Overall pathing
popPathAPUniqMac <- dfGenAP %>% 
  group_by(macaddr, fromAP, toAP, fromLoc, toLoc) %>% 
  summarise(freq = n()) %>% # first group by mac to condense from to info
  group_by(fromAP, toAP, fromLoc, toLoc) %>% 
  summarise(freq = n()) %>% # then group by from, to so that each mac's from to has a weight of 1
  arrange(desc(freq))
  
knitr::kable(popPathAPUniqMac[1:20, ], format = "html")

# Nonadjacent pathing
notAdjAP <- NULL
for(i in 1:500) { # 1:length(popPathAPUniqMac$fromLoc)
  from <- as.character(popPathAPUniqMac$fromLoc[i])
  to <- as.character(popPathAPUniqMac$toLoc[i])
  if(!adjTable[from, to] & from != to) { # if A and B aren't adjacent
    notAdjAP <- rbind(notAdjAP, popPathAPUniqMac[i, ]) # add to dataframe for further analysis
  }
}

knitr::kable(notAdjAP[1:20, ], format = "html")

# Different location jumps
locJumps <- popPathAPUniqMac[which(popPathAPUniqMac$fromLoc != popPathAPUniqMac$toLoc), ] # jumps where locations are different

knitr::kable(locJumps[1:20, ], format = "html")
```



Future work could look at the frequency discrepancies between A -> B and B -> A.
